---
title: "dplyr"
output: pdf_document
date: '2022-08-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## The across function



 


### across

* across() makes it easy to apply the same transformation to multiple columns, allowing you to use select() semantics inside in "data-masking" functions like summarise() and mutate().

* if_any() and if_all() are used to apply the same predicate function to a selection of columns and combine the results into a single logical vector.

* across() supersedes the family of dplyr "scoped variants" like summarise_at(), summarise_if(), and summarise_all() and therefore these functions will not be implemented in poorman.
across: Apply a function (or functions) across multiple columns


### Usage

* across(.cols = everything(), .fns = NULL, ..., .names = NULL)

* if_any(.cols, .fns = NULL, ..., .names = NULL)

* if_all(.cols, .fns = NULL, ..., .names = NULL)

### Arguments

#### .fns
Functions to apply to each of the selected columns. Possible values are:

* NULL, to returns the columns untransformed.

* A function, e.g. mean.

* A lambda, e.g. ~ mean(.x, na.rm = TRUE)

* A list of functions/lambdas, e.g. list(mean = mean, n_miss = ~ sum(is.na(.x))

Within these functions you can use cur_column() and cur_group() to access the current column and grouping keys respectively.

#### ...
Additional arguments for the function calls in .fns.

#### .names
character(n). Currently limited to specifying a vector of names to use for the outputs.

#### cols, .cols
<poor-select> Columns to transform. Because across() is used within functions like summarise() and mutate(), you can't select or compute upon grouping variables.


#### Value

* across() returns a data.frame with one column for each column in .cols and each function in .fns.

* if_any() and if_all() return a logical vector.



### How to use across

There are four columns and I want to quickly get the mean of these columns for each category. First, here’s how I might do this without across:

```{r}
iris  %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length, na.rm = TRUE),
    Sepal.Width = mean(Sepal.Width, na.rm = TRUE),
    Petal.Width = mean(Petal.Width, na.rm = TRUE),
    Petal.Length = mean(Petal.Length, na.rm = TRUE)
    )
```
Which works fine. But imagine if instead of four columns there were 10 or 20 or 100! It would quickly get tedious to add a new line for each column. Here’s where across comes in:

```{r}
iris  %>%
  group_by(Species) %>%
  
  summarise(across(c(Sepal.Length, Sepal.Width, Petal.Length,Petal.Width),mean,na.rm = TRUE))
    
```

Much more efficient. We give across a vector of column names followed by the function (in this case mean) followed by any other arguments we want to apply to the function.





### : 

: for selecting a range of consecutive variables.

```{r}
iris  %>%
  group_by(Species) %>%
       summarise(across(c(Sepal.Length:Petal.Width), mean, na.rm = TRUE))
    
```
### ! 

! for taking the complement of a set of variables.
```{r}
iris  %>%
  group_by(Species) %>%
       summarise(across(!c(Petal.Width), mean, na.rm = TRUE))
    
```

### & and | 

& and |for selecting the intersection or the union of two sets of variables.

```{r}
iris  %>%
  group_by(Species) %>%
       summarise(across(ends_with('Length') & !c(Petal.Length, Petal.Width), mean, na.rm = TRUE))
    
```


### c() 

c() for combining selections.

```{r}
iris  %>%
  group_by(Species) %>%
  
  summarise(across(c(Sepal.Length, Sepal.Width, Petal.Length,Petal.Width),mean,na.rm = TRUE))
    
```



### starts_with()

starts_with(): Starts with a prefix.

```{r}
iris  %>%
  group_by(Species) %>%
  
  summarise(across(starts_with("S"),mean,na.rm = TRUE))
    
```
### ends_with()

ends_with(): Ends with a suffix.

```{r}
iris  %>%
  group_by(Species) %>%
  
  summarise(across(ends_with("dth"),mean,na.rm = TRUE))
    
```
### contains()

contains(): Contains a literal string.

```{r}
iris  %>%
  group_by(Species) %>%
    summarise(across(contains('Length'),mean,na.rm = TRUE))
```

### matches()

matches(): Matches a regular expression.

```{r}
iris  %>%
  group_by(Species) %>%
    summarise(across(matches('^(S|P)'),mean,na.rm = TRUE))
```

### num_range()

num_range(): Matches a numerical range like x01, x02, x03.

```{r}
df <- as.data.frame(matrix(1:24, nrow = 3))
df
df %>% select(num_range("V", seq(1, 1000, by = 3)))
```

```{r}
df <- data.frame(id=c("a","a","b"), tot_1=4:6, tot_2=8:10, tot_3=11:13, tot_4=33:35,tot_5=22:24)
df
df %>% group_by(id) %>%
  mutate(across(.cols = num_range("tot_", seq(1, 5, by = 2)),mean,na.rm = TRUE))
```


```{r}
df %>% group_by(id) %>%
  summarise(across(.cols = num_range(prefix="tot_", range=seq(1, 5, by = 2)),mean,na.rm = TRUE))
```

### all_of() 

all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.






```{r}
iris  %>%
  group_by(Species) %>%
    summarise(across(all_of(c('Sepal.Length','Sepal.Width','Petal.Length')),mean,na.rm = TRUE))
```


### any_of()

any_of(): Same as all_of(), except that no error is thrown for names that don't exist.

```{r}
iris  %>%
  group_by(Species) %>%
    summarise(across(any_of(c('Sepal.Length','Sepal.Width','Petal.Length','Not_valid_name')),mean,na.rm = TRUE))
```

### where()

where(): Applies a function to all variables and selects those for which the function returns TRUE.



```{r}
iris  %>%
  group_by(Species) %>%
       summarise(across(where(is.numeric), mean, na.rm = TRUE))
    
```




## Using in-line functions with across

Let’s look at an example of summarizing the columns using a custom function (rather than n_distinct()). I usually do this using the tilde-dot shorthand for inline functions. The notation works by replacing

```{r}
function(x) {
  x + 10
}
```

with

```{r}
~{.x + 10}

```

 ~ indicates that you have started an anonymous function, and the argument of the anonymous function can be referred to using .x  (or simply  . ). Unlike normal function arguments that can be anything that you like, the tilde-dot function argument is always .x.

For instance, to identify how many missing values there are in every column, we could specify the inline function ~sum(is.na(.)), which calculates how many NA values are in each column (where the column is represented by .) and adds them up:

```{r}
dat<-data.frame(a=c(1,2,3,NA,NA,6),b=1:6,d=c(NA,2:6))
dat
dat %>%
  summarise(across(everything(),  ~sum(is.na(.))))
```


```{r}
dat<-data.frame(a=c(1:4),b=c(1:4)^2,d=c(1:4)^3)
dat
dat %>%
  summarise(across(everything(),  ~ .x +10))
```

























