---
title: "Python Cheat Sheets"
output:
  pdf_document: default
  html_document:
    code_folding: show
    theme:
      bg: '#202123'
      fg: '#B8BCC2'
      primary: '#EA80FC'
      secondary: '#00DAC6'
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
library(reticulate)
#reticulate::use_python("C:/Program Files/Python310/python.exe")
```

## <span style="color:orange"> Pandas Cheat Sheet - part 0001    </span> {.tabset .tabset-pills}


### Choosing a file

#### <span style="color:orange"> Choosing a file in Python with simple Dialog: easygui  </span>
```{python}
#To install:
# pip install easygui
import easygui
#filename =easygui.fileopenbox()
#print(filename)
#easygui.egdemo()

```
#### <span style="color:orange"> Choosing a file in Python with simple Dialog: plyer  </span>
```{python}
#To install:
# pip install plyer
#import plyer 
#filename =plyer.filechooser.open_file()
#print(filename)
```


```{r}
#filename =file.choose()
#print(filename)
```

### read csv and xlsx files

#### <span style="color:orange">  csv </span>

```{python}
import pandas as pd
#import easygui
#filename = easygui.fileopenbox()
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.csv"
df1=pd.read_csv(filename)
df1
```


```{python}
#df1.dtypes
#df1.head()
#df1.tail(2)
#df1.columns
#df1.describe()
#df1["Petal_Length"]
#df1[0:2]
df1[df1.Petal_Length==1.4]
```

#### <span style="color:orange">  xlsx </span>





###  filter and select


#### <span style="color:orange">  filter </span>


```{python}
pl=1.4
qs="Petal_Length==@pl"
df1.query(qs)
```
```{r}
#filename = file.choose()
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.csv"
df1=read.csv(filename)
head(df1)
```

```{python}
pw=.3
sp=["setosa","setosa1"]
qs="Species in @sp"\
 " and Petal_Width <= @pw"
df1.query(qs)
```

```{r}
library(dplyr)
pw=.3
sp=c("setosa","setosa1")
df1 %>%
  dplyr::filter(
    Species %in% sp
    ,Petal_Width <= pw           
  )
```

#### <span style="color:orange">  select </span>

```{python}
cl=["Sepal_Length","Petal_Width"]
df1[cl]
```
```{r}
library(dplyr)
cl=c("Sepal_Length","Petal_Width")
df1%>% dplyr::select(all_of(cl))
# df1%>% dplyr::select(Sepal_Length,Petal_Width)
```

### pivot/melt


#### <span style="color:orange">  melt - pivot_longer </span>



```{python}
import pandas as pd
df2 = pd.DataFrame({'A': {0: 'a', 1: 'b', 2: 'c'},
                   'B': {0: 1, 1: 3, 2: 5},
                   'C': {0: 2, 1: 4, 2: 6}})
df2.melt(id_vars='A')
df2.melt(id_vars='A', value_vars=['B','C'], var_name='BC', value_name='value')
```
```{r}
library(dplyr)
library(tidyr)
df2 = data.frame(
  A=c('a','b','c')
  ,B=c(1,3,5)
  ,C=c(2,4,6)
  )
df2 %>% 
  pivot_longer(B:C,names_to = 'BC',values_to = 'value')
```
```{r}
library(dplyr)
library(tidyr)
df2 = data.frame(
  A=c('a', 'a','b', 'b', 'c','c')
  ,B=c('A', 'B','A', 'B', 'A','B')
  ,D=c( 1, 3, 5,7,9,11)
  ,E=c(2, 4, 6,8,10,12)
  )
df2 %>%
  tidyr::pivot_longer(cols = any_of(c('D','E')),names_to = "DE",values_to  = "value")
```
#### <span style="color:orange"> pivot_wider </span>

```{python}
import pandas as pd
df2 = pd.DataFrame({'A': {0: 'a', 1: 'b', 2: 'c'},
                   'B': {0: 1, 1: 3, 2: 5},
                   'C': {0: 2, 1: 4, 2: 6}})
#print(df2)
df2_melt=df2.melt(id_vars='A', value_vars=['B','C'], var_name='BC', value_name='value')
#print(df2_melt)
df_pivot=df2_melt.pivot(index='A', columns=['BC'])#, values='value')
df2_r = df_pivot.reset_index(None)
df2_r.columns = ['A', 'B', 'C']
print(df2_r)
df2_r.columns=df2.columns.values
print(df2.columns.values)
print(df2_r)
```
```{r}
library(dplyr)
library(tidyr)
df2 = data.frame(
  A=c('a','b','c')
  ,B=c(1,3,5)
  ,C=c(2,4,6)
)
df2_melt<-df2 %>%
  tidyr::pivot_longer(cols = any_of(c('B','C')),names_to = "BC",values_to  = "value")
df_pivot <- df2_melt %>%
  tidyr::pivot_wider(id_cols = A, names_from = BC,values_from = value )
df_pivot
```



### dplython

```{python}
#To install:
# pip install dplython
```

#### <span style="color:orange">  1 </span>

### siuba

```{python}
#To install:
# pip install siuba
```

#### <span style="color:orange">  1 </span>








## Contact us

Contact me at masoudfaridi@modares.ac.ir or masoud1faridi@gmail.com

