---
title: "RPandas Part 001"
output: pdf_document
date: '2023.03.22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(dplyr)
#tinytex::install_tinytex()
```

## Choosing a file

### <span style="color:orange"> Choosing a file in Python with simple Dialog: easygui  </span>
```{python}
# python code
#To install:
# pip install easygui
import easygui
#filename =easygui.fileopenbox()
#print(filename)
#easygui.egdemo()

```


### <span style="color:orange"> Choosing a file in Python with simple Dialog: plyer  </span>
```{python}
# python code
#To install:
# pip install plyer
#import plyer 
#filename =plyer.filechooser.open_file()
#print(filename)
```


```{r}
# R code
#filename =file.choose()
#print(filename)
```

## read csv and xlsx files

### <span style="color:orange">  csv </span>

```{python}
# python code
import pandas as pd
#import easygui
#filename = easygui.fileopenbox()
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.csv"
df1=pd.read_csv(filename)
df1
```


```{python} 
# python code
#df1.dtypes
#df1.head()
#df1.tail(2)
#df1.columns
#df1.describe()
#df1["Petal_Length"]
#df1[0:2]
df1[df1.Petal_Length==1.4]
```
```{r}
# R code
#filename = file.choose()
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.csv"
df1=read.csv(filename)
head(df1)
```


### <span style="color:orange">  xlsx </span>

#### R Code

```{r}
#R code
library(openxlsx)
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.xlsx"
df1=openxlsx::read.xlsx(filename)
head(df1)
```
```{r}
#R code
#read all sheets
library(openxlsx)
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.xlsx"
SheetNames <- openxlsx::getSheetNames(filename)
SheetNames
SheetList <- lapply(SheetNames,openxlsx::read.xlsx,xlsxFile=filename)
names(SheetList) <- SheetNames
SheetList$Sheet1[1:4,]
SheetList$iris[1:4,]

```
```{r}
# write xlsx files
library(openxlsx)
wb <- createWorkbook()  #wb <- loadWorkbook("RawExcel.xlsx")
addWorksheet(wb, sheetName = "sheetname1")
writeData(wb, sheet = "sheetname1", x = SheetList$iris[1:4,])
addWorksheet(wb, sheetName = "sheetname2")
writeData(wb, sheet = "sheetname2", x = SheetList$Sheet1[1:4,])
#saveWorkbook(wb, "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris2.xlsx")

```


#### Python Code

Python Code
```{python}
import pandas as pd
xls = pd.ExcelFile('G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris2.xlsx')
xls.sheet_names
df1 = pd.read_excel(xls, xls.sheet_names[0])
df2 = pd.read_excel(xls, xls.sheet_names[1])
df1
```
```{python}
import pandas as pd
xls = pd.ExcelFile('G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris2.xlsx')
xls.sheet_names
df1 = pd.read_excel(xls, xls.sheet_names[0])
df2 = pd.read_excel(xls, xls.sheet_names[1])
df1
dff=[pd.read_excel(xls, x) for x in xls.sheet_names]

```
```{python}
import pandas as pd
dict_temp = pd.read_excel('G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris2.xlsx', sheet_name= xls.sheet_names)
dict_temp['sheetname1']
dict_temp['sheetname2']
```




##  filter and select


### <span style="color:orange">  filter </span>


```{python} 
# python code
pl=1.4
qs="Petal_Length==@pl"
df1.query(qs)
```

```{python} 
# python code
pw=.3
sp=["setosa","setosa1"]
qs="Species in @sp"\
 " and Petal_Width <= @pw"
df1.query(qs)
```

```{r}
# R code
filename = "G:\\Python tutorial\\pythontutorial\\pythontutorial\\iris.csv"
df1=read.csv(filename)
pw=.3
sp=c("setosa","setosa1")
df1 %>%
  dplyr::filter(
    Species %in% sp
    ,Petal_Width <= pw           
  ) %>% head()
```

### <span style="color:orange">  select </span>

```{python} 
# python code
cl=["Sepal_Length","Petal_Width"]
df1[cl]
```
```{r}
# R code
cl=c("Sepal_Length","Petal_Width")
df1%>% dplyr::select(all_of(cl))  %>% head()
# df1%>% dplyr::select(Sepal_Length,Petal_Width)
```


###  starts_with


```{r}
# R Code
names(iris)
iris %>%
   head(3)
iris %>%
  dplyr::select(starts_with('Sepal')) %>% head(3)
```

###  ends_with


```{r}
# R Code
iris %>%
  dplyr::select(ends_with('Length')) %>% head(3)
```


###  contains


```{r}
# R Code
iris %>%
  dplyr::select(contains('Wid')) %>% head(3)
```


###  matches


```{r}
# R Code
iris %>%
  dplyr::select(dplyr::matches("^(S)")) %>% head(3)
iris %>%
  dplyr::select(dplyr::matches("^(Sepal.L|Sp)")) %>% head(3)
```



###  any_of


```{r}
# R Code
iris %>%
  dplyr::select(dplyr::any_of(c('Sepal.Length','Sepal.Width','Petal.Length','Not_valid_name'))) %>% head(3)

```




###  where


```{r}
# R Code
iris %>%
  dplyr::select(dplyr::where(is.numeric)) %>% head(3)
```


## siuba python



```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# old approach: repeat name
cars[cars.cyl == 4]
# old approach: lambda
cars[lambda _: _.cyl == 4]
# siu approach
cars[_.cyl == 4]
```

### filter







[siuba.org](https://siuba.org/)


```{python}
import siuba
from siuba.data import cars
df2=(cars
    >> siuba.group_by(siuba._.cyl)
    >> siuba.filter(siuba._.mpg < siuba._.mpg.mean())
)
print(df2)
```
```{python}
import siuba
from siuba.data import cars
from siuba import _
df2=(cars
    >> siuba.group_by(_.cyl)
    >> siuba.filter(_.mpg < _.mpg.mean())
)
print(df2)
```

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df2=(cars
    >> siuba.group_by( _.cyl )
    >> siuba.filter( _.mpg < _.mpg.mean() )
)
print(df2.head())

```


### isin (%in%)

```{python}
import siuba
from siuba.data import cars
from siuba import _
import pandas as pd
df2=(cars
    >> siuba.filter( _.cyl.isin([8,6]))
)
print( df2)
```
```{python}
import siuba
from siuba.data import cars
from siuba import _
import pandas as pd
df2=(cars
    >> siuba.filter( _.cyl.isin([8,6]), _.hp.isin([123,180]))
)
print( df2)
```


```{python}
import siuba
from siuba.data import cars
from siuba import _
import pandas as pd
df3 = pd.DataFrame({
  'Country': ['Germany', 'France', 'Spain', 'Portugal']
  ,'num':[101,209,200,100]
  })
df4=(df3 >>
 siuba.filter(_.Country.isin(['Germany','Portugal']))
)
print(df4)
```




### select

```{python}
import siuba
from siuba.data import cars
from siuba import _
df2=(cars
    >> siuba.select(_.mpg , _.cyl )
)
print( df2.head() )
```
### Select by name or position




```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _

df2=(cars
    >> siuba.select(_.cyl, 1, ~_.hp )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _

df2=(cars
    >> siuba.select( -1,-2 )
)
print( df2.head() )
```




```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _

df2=(cars
    >> siuba.select( ~_.mpg )
)
print( df2.head() )
```

### Renaming columns

You can rename a specified column by using the equality operator (==). This operation takes the following form.


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
#_.new_name == _.old_name
df2=(cars
    >> siuba.select( _.hp2==_.hp, _.cyl2==_.cyl )
)
print( df2.head() )
```

### Use indexing (e.g. _["some_name"]) to refer to any column by name.

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_["start_name":"end_name"])
df2=(cars
    >> siuba.select(_['mpg' , 'hp'] )
)
print( df2.head() )
```



### Select by slice _[start:end]


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_["start_name":"end_name"])
df2=(cars
    >> siuba.select(_.cyl,_['mpg' : 'hp'] )
)
print( df2.head() )
```

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_["start_name":"end_name"])
df2=(cars
    >> siuba.select(_.cyl,_[1:3])
)
print( df2.head() )
```



### Exclusion

You can exclude slice selections using the ~ operator.


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(~_["start_name":"end_name"])
df2=(cars
    >> siuba.select(_.cyl,~_['mpg' : 'hp'] )
)
print( df2.head() )
```

### Select by pattern (e.g. endswith)


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.endswith("mm"))
df2=(cars
    >> siuba.select(_.cyl,_.endswith('p') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.endswith("mm"))
df2=(cars
    >> siuba.select(~_.endswith('p') )
)
print( df2.head() )
```

### Select by pattern (e.g. startswith)


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.startswith("mm"))
df2=(cars
    >> siuba.select(_.cyl,_.startswith('m') )
)
print( df2.head() )
```

### Select by pattern (e.g. contains)


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(_.cyl,_.contains('p') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(_.contains('h|c') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(~_.contains('h|c') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(_.contains('^c|g$') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(_.contains('.{2}g$') )
)
print( df2.head() )
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.select(_.contains('p.+') )
)
print( df2.head() )
```

### mutate 



```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.mutate(mpg_std=(_.mpg-_.mpg.mean())/_.mpg.std(),mpg_hp=_.mpg*_.hp,hp_per_cyl = _.hp / _.cyl)
)
print( df2.head() )
```

### Grouped mutates


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df5=(cars
  >> siuba.group_by( _.cyl )
  >> siuba.mutate(
       hp_mean = _.hp.mean(),
       demeaned_hp = _.hp - _.hp_mean
     )
  )
print(df5)
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','B','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=(df6
  >> siuba.group_by( _.g1 )
  >> siuba.mutate(
       vl_mean = _.vl.mean(),
       vl_std = _.vl.std()
       ,vl_z=(_.vl-_.vl.mean())/_.vl.std()
     )
  )
print(df5)
```
```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','B','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=(df6
  >> siuba.group_by( _.g1 ,_.g2)
  >> siuba.mutate(
       vl_mean = _.vl.mean(),
       vl_std = _.vl.std()
       ,vl_z=(_.vl-_.vl.mean())/_.vl.std()
     )
  )
print(df5)
```


### shift



```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','B','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=df5=(df6
  >> siuba.mutate(
       vl_lag = _.vl - _.vl.shift(1)
     )
  )
print(df5)
```

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','B','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=df5=(df6
  >> siuba.mutate(
       vl_lag = _.vl - _.vl.shift(-1)
     )
  )
print(df5)
```


### Grouped shifts


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=df5=(df6
  >> siuba.group_by(_.g1)
  >> siuba.mutate(
       vl_lag = _.vl - _.vl.shift(1)
     )
  )
print(df5)
```





### summarize



Summarize over all rows

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
# select(_.species, _.contains("mm"))
df2=(cars
    >> siuba.summarize(mpg_sum=_.mpg.sum(),mpg_mean=_.mpg.mean(),mpg_min=_.mpg.min(),mpg_max=_.mpg.max()
     , mpg_std=_.mpg.std(), mpg_median=_.mpg.median(),mpg_count=_.mpg.count())
)
print( df2.head() )
```

### Summarize over groups


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','B','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=(df6
  >> siuba.group_by( _.g1 )
  >> siuba.summarize(
       vl_mean = _.vl.mean()
       ,vl_std=_.vl.std()
       ,vl_min=_.vl.min()
       ,vl_max=_.vl.max()
       ,vl_median=_.vl.median()
     )
  )
print(df5)
```

```{python}
import pandas as pd
import siuba
from siuba.data import mtcars
from siuba import _
df5=(mtcars
  >> siuba.group_by(_.cyl)
  >> siuba.summarize(
       avg = _.mpg.mean(),
       range = _.mpg.max() - _.mpg.min(),
       avg_per_cyl = (_.mpg / _.cyl).mean()
  )
)
print(df5)
```


### Group by an expression


```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
    ,'vl':[1,2,3,7,8,9]
 }
 )
df5=(df6
  >> siuba.group_by( g1_temp= _.vl>3 )
  >> siuba.summarize(
       vl_mean = _.vl.mean()
       ,vl_std=_.vl.std()
       ,vl_min=_.vl.min()
       ,vl_max=_.vl.max()
       ,vl_median=_.vl.median()
     )
  )
print(df5)
```
group_by(high_hp = _.hp > 300)


### Count rows

```{python}
import pandas as pd
import siuba
from siuba.data import cars
from siuba import _
df6=pd.DataFrame(
 {
   'g1':['a','a','a','b','b','b']
   ,'g2':['A','A','B','A','A','B']
   ,'vl':[1,2,3,7,8,9]
 }
 )
df5=(df6
  >> siuba.group_by( _.g1 )
  >> siuba.summarize(
      count=_.shape[0]
     )
  )
print(df5)
df5=(df6
  >> siuba.group_by( _.g2 )
  >> siuba.summarize(
      count=_.shape[0]
     )
  )
print(df5)
```


### sort_values


```{python}
from siuba import _, count
from siuba.data import mtcars

df3=(mtcars
    >> count(_.cyl)         # this is a siuba verb
    >> _.sort_values("n")   # this is a pandas method
)
print(df3)
```



### shape

```{python}
from siuba import _, count
from siuba.data import mtcars

df3=(mtcars
    >> _.shape
)
print(df3)

df3=(mtcars
    >> _.shape[0]
)
print(df3)

df3=(mtcars
    >> _.shape[1]
)
print(df3)
```

### Google colab overrides _ (import _ as X)


```{python}
from siuba import _ as XX, filter
from siuba.data import mtcars

df3=mtcars >> filter(XX.mpg > 30)
print(df3)
```

### Dropping NA values

Use pandas.isna() / pandas.notna() to determine whether a value is considered to be NA.



```{python}
df4 = pd.DataFrame({
    "x": [True, False, None,True, False, None],
    "value":[0,5,6,3,9,8]
    })

print(df4)
```
```{python}
print(pd.isna(df4.x))
print(df4[pd.isna(df4.x)])
print(df4[pd.notna(df4.x)])
```
```{python}
import pandas as pd
from siuba import _, count
from siuba.data import mtcars

df5=df4 >> siuba.filter(_.x)
print(df5)
```


```{python}
import pandas as pd
from siuba import _, count
from siuba.data import mtcars

df5=df4 >> siuba.filter(_.x.notna())
print(df5)
```
```{python}
import pandas as pd
from siuba import _, count
from siuba.data import mtcars

df5=df4 >> siuba.filter(_.x.notna(),_.value)
print(df5)
```
```{python}
import pandas as pd
from siuba import _, count
from siuba.data import mtcars

df5=df4 >> siuba.filter(_.x,_.value)
print(df5)
```


### arrange


```{python}
import pandas as pd
import siuba
from siuba import _, count
from siuba.data import cars

df5= cars >> siuba.arrange(-_.hp)
print(df5.head())
```

```{python}
import pandas as pd
import siuba
from siuba import _, count
from siuba.data import cars

df5= cars >> siuba.arrange(_.cyl,-_.mpg)
print(df5)
df5=cars >> siuba.arrange(_.hp / _.cyl)
print(df5)
```


### Using with plotnine

```{python}
from siuba import mutate, _
from siuba.data import mtcars
from plotnine import ggplot, aes, geom_point

(mtcars
  >> mutate(hp_per_cyl = _.hp / _.cyl)
  >> ggplot(aes("cyl", "hp_per_cyl"))
   + geom_point()
)
```




### Call external functions

```{python}
import pandas as pd
from siuba import _, mutate
from siuba.siu import call

my_dates = pd.DataFrame({"date": ["2021-01-01", "2021-01-02"]})

pd.to_datetime(my_dates.date)
```



```{python}
my_dates >> mutate(parsed = _.date) >> _.parsed
```

```{python}
my_dates >> mutate(parsed = call(pd.to_datetime, _.date))
```

### sql basic

[https://siuba.org/guide/basics-sql.html](https://siuba.org/guide/basics-sql.html)


## pivot/melt


### <span style="color:orange">  melt - pivot_longer </span>



```{python} 
# python code
import pandas as pd
df2 = pd.DataFrame({'A': {0: 'a', 1: 'b', 2: 'c'},
                   'B': {0: 1, 1: 3, 2: 5},
                   'C': {0: 2, 1: 4, 2: 6}})
df2.melt(id_vars='A')
df2.melt(id_vars='A', value_vars=['B','C'], var_name='BC', value_name='value')
```
```{r}
# R code
library(tidyr)
df2 = data.frame(
  A=c('a','b','c')
  ,B=c(1,3,5)
  ,C=c(2,4,6)
  )
df2 %>% 
  pivot_longer(B:C,names_to = 'BC',values_to = 'value')  %>% head()
```
```{r}
# R code
library(tidyr)
df2 = data.frame(
  A=c('a', 'a','b', 'b', 'c','c')
  ,B=c('A', 'B','A', 'B', 'A','B')
  ,D=c( 1, 3, 5,7,9,11)
  ,E=c(2, 4, 6,8,10,12)
  )
df2 %>%
  tidyr::pivot_longer(cols = any_of(c('D','E')),names_to = "DE",values_to  = "value")  %>% head()
```
### <span style="color:orange"> pivot_wider </span>

```{python}
# python code
import pandas as pd
df2 = pd.DataFrame({'A': {0: 'a', 1: 'b', 2: 'c'},
                   'B': {0: 1, 1: 3, 2: 5},
                   'C': {0: 2, 1: 4, 2: 6}})
#print(df2)
df2_melt=df2.melt(id_vars='A', value_vars=['B','C'], var_name='BC', value_name='value')
#print(df2_melt)
df_pivot=df2_melt.pivot(index='A', columns=['BC'])#, values='value')
df2_r = df_pivot.reset_index(None)
df2_r.columns = ['A', 'B', 'C']
print(df2_r)
df2_r.columns=df2.columns.values
print(df2.columns.values)
print(df2_r)
```
```{r}
# R code
library(tidyr)
df2 = data.frame(
  A=c('a','b','c')
  ,B=c(1,3,5)
  ,C=c(2,4,6)
)
df2_melt<-df2 %>%
  tidyr::pivot_longer(cols = any_of(c('B','C')),names_to = "BC",values_to  = "value")
df_pivot <- df2_melt %>%
  tidyr::pivot_wider(id_cols = A, names_from = BC,values_from = value )
df_pivot  %>% head()
```
















## The across function 


### across

* across() makes it easy to apply the same transformation to multiple columns, allowing you to use select() semantics inside in "data-masking" functions like summarise() and mutate().

* if_any() and if_all() are used to apply the same predicate function to a selection of columns and combine the results into a single logical vector.

* across() supersedes the family of dplyr "scoped variants" like summarise_at(), summarise_if(), and summarise_all() and therefore these functions will not be implemented in poorman.
across: Apply a function (or functions) across multiple columns


### Usage

* across(.cols = everything(), .fns = NULL, ..., .names = NULL)

* if_any(.cols, .fns = NULL, ..., .names = NULL)

* if_all(.cols, .fns = NULL, ..., .names = NULL)

### Arguments

#### .fns
Functions to apply to each of the selected columns. Possible values are:

* NULL, to returns the columns untransformed.

* A function, e.g. mean.

* A lambda, e.g. ~ mean(.x, na.rm = TRUE)

* A list of functions/lambdas, e.g. list(mean = mean, n_miss = ~ sum(is.na(.x))

Within these functions you can use cur_column() and cur_group() to access the current column and grouping keys respectively.

#### ...
Additional arguments for the function calls in .fns.

#### .names
character(n). Currently limited to specifying a vector of names to use for the outputs.

#### cols, .cols
<poor-select> Columns to transform. Because across() is used within functions like summarise() and mutate(), you can't select or compute upon grouping variables.


#### Value

* across() returns a data.frame with one column for each column in .cols and each function in .fns.

* if_any() and if_all() return a logical vector.



### How to use across

There are four columns and I want to quickly get the mean of these columns for each category. First, here’s how I might do this without across:

```{r}
# R code
iris  %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length, na.rm = TRUE),
    Sepal.Width = mean(Sepal.Width, na.rm = TRUE),
    Petal.Width = mean(Petal.Width, na.rm = TRUE),
    Petal.Length = mean(Petal.Length, na.rm = TRUE)
    )  %>% head()
```
Which works fine. But imagine if instead of four columns there were 10 or 20 or 100! It would quickly get tedious to add a new line for each column. Here’s where across comes in:

```{r}
# R code
iris  %>%
  group_by(Species) %>%
  
  summarise(across(c(Sepal.Length, Sepal.Width, Petal.Length,Petal.Width),mean,na.rm = TRUE)) %>% head()
    
```

Much more efficient. We give across a vector of column names followed by the function (in this case mean) followed by any other arguments we want to apply to the function.





### : 

: for selecting a range of consecutive variables.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
       summarise(across(c(Sepal.Length:Petal.Width), mean, na.rm = TRUE))  %>% head()
    
```
### ! 

! for taking the complement of a set of variables.
```{r}
# R code
iris  %>%
  group_by(Species) %>%
       summarise(across(!c(Petal.Width), mean, na.rm = TRUE))  %>% head()
    
```

### & and | 

& and |for selecting the intersection or the union of two sets of variables.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
       summarise(across(ends_with('Length') & !c(Petal.Length, Petal.Width), mean, na.rm = TRUE))  %>% head()
    
```


### c() 

c() for combining selections.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
  
  summarise(across(c(Sepal.Length, Sepal.Width, Petal.Length,Petal.Width),mean,na.rm = TRUE)) %>% head()
    
```



### starts_with()

starts_with(): Starts with a prefix.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
   
  summarise(across(starts_with("S"),mean,na.rm = TRUE))  %>% head()
    
```
### ends_with()

ends_with(): Ends with a suffix.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
  
  summarise(across(ends_with("dth"),mean,na.rm = TRUE))  %>% head()
    
```
### contains()

contains(): Contains a literal string.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
    summarise(across(contains('Length'),mean,na.rm = TRUE))  %>% head()
```

### matches()

matches(): Matches a regular expression.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
    summarise(across(matches('^(S|P)'),mean,na.rm = TRUE))  %>% head()
```

### num_range()

num_range(): Matches a numerical range like x01, x02, x03.

```{r}
# R code
df <- as.data.frame(matrix(1:24, nrow = 3))
df  %>% head()
df %>% select(num_range("V", seq(1, 1000, by = 3)))  %>% head()
```

```{r}
# R code
df <- data.frame(id=c("a","a","b"), tot_1=4:6, tot_2=8:10, tot_3=11:13, tot_4=33:35,tot_5=22:24)
df  %>% head()
df %>% group_by(id) %>%
  mutate(across(.cols = num_range("tot_", seq(1, 5, by = 2)),mean,na.rm = TRUE))  %>% head()
```


```{r}
# R code
df %>% group_by(id) %>%
  summarise(across(.cols = num_range(prefix="tot_", range=seq(1, 5, by = 2)),mean,na.rm = TRUE))  %>% head()
```

### all_of() 

all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.






```{r}
# R code
iris  %>%
  group_by(Species) %>%
    summarise(across(all_of(c('Sepal.Length','Sepal.Width','Petal.Length')),mean,na.rm = TRUE))  %>% head()
```


### any_of()

any_of(): Same as all_of(), except that no error is thrown for names that don't exist.

```{r}
# R code
iris  %>%
  group_by(Species) %>%
    summarise(across(any_of(c('Sepal.Length','Sepal.Width','Petal.Length','Not_valid_name')),mean,na.rm = TRUE))  %>% head()
```

### where()

where(): Applies a function to all variables and selects those for which the function returns TRUE.



```{r}
# R code
iris  %>%
  group_by(Species) %>%
       summarise(across(where(is.numeric), mean, na.rm = TRUE))  %>% head()
    
```




## Using in-line functions with across

Let’s look at an example of summarizing the columns using a custom function (rather than n_distinct()). I usually do this using the tilde-dot shorthand for inline functions. The notation works by replacing

```{r}
# R code
function(x) {
  x + 10
}
```

with

```{r}
# R code
~{.x + 10}

```

 ~ indicates that you have started an anonymous function, and the argument of the anonymous function can be referred to using .x  (or simply  . ). Unlike normal function arguments that can be anything that you like, the tilde-dot function argument is always .x.

For instance, to identify how many missing values there are in every column, we could specify the inline function ~sum(is.na(.)), which calculates how many NA values are in each column (where the column is represented by .) and adds them up:

```{r}
# R code
dat<-data.frame(a=c(1,2,3,NA,NA,6),b=1:6,d=c(NA,2:6))
dat
dat %>%
  summarise(across(everything(),  ~sum(is.na(.))))  %>% head()
```


```{r}
# R code
dat<-data.frame(a=c(1:4),b=c(1:4)^2,d=c(1:4)^3)
dat  %>% head()
dat %>%
  summarise(across(everything(),  ~ .x +10))  %>% head()
```



## Contact us

Contact me at masoudfaridi@modares.ac.ir or masoud1faridi@gmail.com





















